---
title: "Homework 4"
author: "LeTicia Cancel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


You get to decide which dataset you want to work on. The data set must be different from the ones used in previous homework You can work on a problem from your job, or something you are interested in. You may also obtain a dataset from sites such as Kaggle, Data.Gov, Census Bureau, USGS or other open data portals. 

Select one of the methodologies studied in weeks 1-10, and one methodology from weeks 11-15 to apply in the new dataset selected. To complete this task:. 

- describe the problem you are trying to solve.
- describe your datases and what you did to prepare the data for analysis. 
- methodologies you used for analyzing the data
- what's the purpose of the analysis performed
- make your conclusions from your analysis. Please be sure to address the business impact (it could be of any domain) of your solution.

Your final presentation could be the traditional R file or Python file and essay, or it could be an oral presentation with the execution and explanation of your code, recorded on any platform of your choice (Youtube, Free Cam). If you select the presentation, it should be a 5 to 8 minutes recording. 


```{r}
#libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(rpart)
library(rpart.plot)
library(rcompanion)
library(datasets)
library(caTools)
library(party)
library(caret)
library(ggcorrplot)
library(randomForest)
library(e1071)
```


## Data Import

```{r}
bank_df <- read.csv("https://raw.githubusercontent.com/ltcancel/DATA622/main/Homework4/Data/Bank_Personal_Loan_Modelling.csv")
?read.table

head(bank_df)
```

## Data Exploration

Exploring each dataframe

```{r fig.width=20, fig.height=10}
summary(bank_df)
```

Convert Bit columns to factor

```{r fig.height=12, fig.width=20}
bank_df$Education <- as.factor(bank_df$Education)
bank_df$Personal.Loan <- as.factor(bank_df$Personal.Loan)
bank_df$Securities.Account <- as.factor(bank_df$Securities.Account)
bank_df$CD.Account <- as.factor(bank_df$CD.Account)
bank_df$Online <- as.factor(bank_df$Online)
bank_df$CreditCard <- as.factor(bank_df$CreditCard)
```

Column Exploration

Graphs of categorical variables
```{r}
hist(bank_df$Family, main = "Family Size of Customer", col = "orange")
plot(bank_df$Education, main = "Educational Level of Customer", col = "orange")
plot(bank_df$Personal.Loan, main = "Customers Accepting Personal Loan Offer", col = "orange")
plot(bank_df$Securities.Account, main = "Security Account Holder", col = "orange")
plot(bank_df$CD.Account, main = "CD Account Holder", col = "orange")
plot(bank_df$Online, main = "Internet Banking", col = "orange")
plot(bank_df$CreditCard, main = "Owns a Credit Card by UniveralBank", col = "orange")

```

Boxplots of continuous variables

```{r}
boxplot(bank_df$Age, main = "Customer Age")
boxplot(bank_df$Experience, main = "Customer Years of Professional Experience")
boxplot(bank_df$Income, main = "Customer Annual Income")
boxplot(bank_df$CCAvg, main = "Customer Credit Card Average Monthly Spending")
boxplot(bank_df$Mortgage, main = "Customer Home Mortgage")

```

Decision Tree

Create the train/test sets
```{r}
set.seed(123)
create_train_test <- function(data, size = 0.8, train = TRUE){
  n_row = nrow(data)
  total_row = size * n_row
  train_sample <- 1: total_row
  if (train == TRUE){
    return(data[train_sample, ])
  } else {
    return(data[-train_sample, ])
  }
}
```


Test the function and check the dimension
```{r}
data_train <- create_train_test(bank_df, 0.8, train = TRUE)
data_test <- create_train_test(bank_df, 0.8, train = FALSE)
dim(data_train)
dim(data_test)
```

Decision Tree using Personal Loan as the predictor
```{r fig.height=15, fig.width=20}
control <- rpart.control(minsplit = 5L, maxdepth = 5L, minbucket = 5, cp = 0.002, maxsurrogate = 4)
fit <- rpart(Personal.Loan~., data = data_train, method = 'class', control = control)
rpart.plot(fit, extra = "auto")
```



Prune the tree

```{r}
printcp(fit)
plotcp(fit)
```


Confusion Matrix

```{r}
predict_df <- predict(fit, data_test, type = "class")


#?predict
print("Decision Tree Confusion Matrix")
confusionMatrix(predict_df, data_test$Personal.Loan)
```






